\section{Regular expressions and finite automatons}
\label{sec:regular_finite}
%% \todo[inline]{MAn bruger dem til at beskrive tekststrenge med bestemte
%% mønstre, prøv at sætte noget lettere fordøjeligt ind før definitionerne}

A regular language is a possibly infinite set of finite sequences of
symbols from a finite alphabet. It is a formal language that must
fulfill a number of properties.

%\todo{wikipedia havde noget interessant at sige men vi kan ikke bare
%  kopiere det ind :(}

We provide this formal definition of regular languages:
\begin{definition}[Regular language]
  The regular language over the alphabet $\Sigma$ is defined
  recursively as:
  \begin{itemize}
  \item The empty language $\emptyset$.
  \item The empty string language $\{\upvarepsilon\}$.
  \item The singleton language $\{a\}$, for any symbol $a\in \Sigma$.
  \item If $L_r$ and $L_s$ are both regular languages then the union
    $L_r \cup L_s$ is also a regular language.
  \item If $L_r$ and $L_s$ are both regular languages then the
    concatenation $L_r \bullet L_s$ is also a regular language.
  \item if $L$ is a regular language then the Kleene star\footnote{We use the conventional definition of the Kleene star; a unary operator meaning ``zero or more'' } $L*$ is also
    a regular language.
  \end{itemize}
\end{definition}



\subsection{Regular expressions}

Regular expressions are used to denote regular languages. They are
written in a formal language consisting of two types of characters:
Meta characters and literal characters. The meta characters have
special meaning and are interpreted by a regular expression
engine. Some of the basic meta characters include parenthesis, the alternation
operator and the Kleene star. Parenthesis provides grouping,
alternation allows the choice between different text strings and the
Kleene star repeats. 

The literal characters have no special meaning; they simply match literally.

%% \begin{definition}[Regular language]
%% A regular language over the alphabet $\Sigma$ can be defined
%% recursively as follows:
%% \begin{description}
%% \item[Basis clauses] 
%%   \begin{itemize}
%%   \item The empty language $\emptyset$
%%   \item The empty string language $\{\upvarepsilon\}$
%%   \item The singleton language $\{a\}$, for any symbol $a\in \Sigma$
%%   \end{itemize}
%%   are all regular languages
%% \item[Inductive clauses]
%%   If $L_r$ and $L_s$ are both regular languages then
%%   \begin{itemize}
%%   \item The union $L_r \cup L_s$
%%   \item The concatenation $L_r \bullet L_s$
%%   \item The kleene star $L_r*$ 
%%   \end{itemize}
%%   are all regular languages
%% \item[Extremal clause] No other languages over $\Sigma$ are regular
%% \end{description}
%% \end{definition}

%% Regular expressions are used to regular languages:

A formal definition of regular expressions:
\begin{definition}[Regular expression]
  \label{def:regexp}
  A regular expression over an alphabet $\Sigma$ can be defined as follows:
\begin{itemize}
\item An empty string, $\upvarepsilon$, and any character from the
  alphabet $\Sigma$
\item If $r_1$ and $r_2$ are regular expressions, then the
  concatenation $r_1r_2$ are also a regular expression
\item If $r_1$ and $r_2$ are regular expressions, then the alternation
  $r_1|r_2$ is also a regular expression
\item If $r$ is a regular expression, then so is the repetition $r*$
\end{itemize}

Any expression is a regular expression if it follows from a finite
number of applications of the above rules.
\end{definition}

The precedence of the operators are: repetition, concatenation and
alternation, from highest to lowest. Concatenation and alternation are
both left-associative. 

\begin{example}[Regular expression]
  Here we have a somewhat complicated example of a regular expression
  that demonstrates the basic operators.
\label{ex:regular1}
Consider the sentence:
\begin{quote}
\textsl{This book was written using 100\% recycled
  words.}\footnote{Terry Pratchett, Wyrd Sisters}
\end{quote}

Other writings such as papers and novels also use words. If we
want to catch sentences referring to these writings as well, we can
use the regular expression: \texttt{(book|paper|novel)}. 

To match the number 100 in the sentence, we could use the regular
expression \texttt{100}. In most cases however we will not know
beforehand how many words are recycled, so we may want to use the
regular expression \texttt{(0|1|2|3|4|5|6|7|8|9)*}, which will match
any natural number.

With this in mind we can write a regular expression to match our
sentence:
\begin{quote}
  \texttt{This (book|paper|novel) was written using \\
    (0|1|2|3|4|5|6|7|8|9)*\% recycled words.}
\end{quote}

\end{example}

\subsection{Extensions to the regular expressions}

Many tools extend the regular expressions presented in the previous
section. A typical extensions is new notation to make it easier to specify
patterns. In this section we present the extensions to definition
\vref{def:regexp} we have made: Additional quantifiers, character classes,
a quoting symbol, a wild card and non-capturing parenthesis.

\begin{itemize}
\item The quantifier $+$ causes the regular expression $r$ to be
  matched one or more times. This can also be written as $rr*$
\item The quantifier $?$ causes the regular expression $r$ to be
  matched zero or one times. This can also be written as
  $\upvarepsilon|r$
\item A character class is delimited by $[]$ and matches exactly one
  character in the input string. Special characters loose their
  meaning inside a character class; $*$, $+$, $?$, $($, $)$ and so on
  are treated as literals.

  Characters can be listed individually, e.g. \texttt{[abc]}, or they
  can be listed as ranges with the range operator: $-$,
  e.g. \texttt{[a-z]}. These can be rewritten in terms of our original
  regular expression: \texttt{a|b|c} and \texttt{a|b|c...x|y|z}
  respectively.

  To match characters not within the range, the complement operator is
  used. \texttt{\^{}} used as the first character in a character
  class, elsewhere it will simply match literally, indicates that only
  characters not listed in the character class should
  match. E.g. \texttt{[\^{}\^{}]} will match anything but a
  \texttt{\^{}}
\item The quoting character \textsf{\textbackslash} will allow the
  operators to match literally. We use \textsf{\textbackslash *} to
  match a \textsl{*}.
\item The wild card \texttt{.} will match any character, including a
  newline.
\item For the non-capturing parenthesis we have the choice of
  notation. Here we will list some of the options, where $r$ is some
  regular expression:
  \begin{itemize}
  \item The industry standard, to which Perl, Python, RE2 and most
    others adhere is: $(?:r)$.
  \item Perl 6 \cite{Wall2002} suggests use of square parenthesis instead:
    $[r]$. These are however already in use by the character classes.
  \item A more intuitive notation could be using single parenthesis
    for non-capturing, $(r)$, and double parenthesis for capturing,
    $((r))$.
  \item A currently unused option is $\{r\}$ as special notation,
    which would be simple to implement. This is however the industry
    standard for repetition notation.
  \end{itemize}
  
  Since there is a standard, we will adhere to it, and use $(?:r)$ for
  non-capturing parenthesis.
\end{itemize}
\begin{example}

As we saw in example \ref{ex:regular1}, we can match a natural number
with the regular expression \texttt{(0|1|2|3|4|5|6|7|8|9)*}. Using the
expansions to regular expressions above, we can rewrite this as:
\begin{description}
\item[\texttt{[0-9]*}] This literally means the same thing.
\item[\texttt{[0-9]+}] We can use a different repetition operator and
  require there be at least one digit.
\item[\texttt{[1-9][0-9]*}] This matches any natural number as
  well, but it will not match any preceding zeros. This is a
  refinement, in that it will match fewer text strings than the first
  expression. It is up to the expression writer to decide what the desired outcome is.
\end{description}
\end{example}


\subsection{Finite automatons}

Finite automatons are used to solve a wide array of problems. In this
thesis we will focus on finite automatons as they are used with
regular expressions. 
A finite state machine consists of a number of
states and transitions between states. It is constructed as follows:
\begin{itemize}
    \item One state is marked as the
    initial state
    \item a set of states, one or more, is marked as final
    \item A condition is attached to each transition between states
    \item Input is consumed in sequence and for each symbol transitions are taken when their attached condition are met
    \item If the simulation ends in a final state, the finite automaton is said to accept the input
\end{itemize}

Finite automatons can be divided in two main categories: The
deterministic (DFA) and the non-deterministic (NFA) finite
automaton. This distinction is mostly relevant in practice, as they
are equivalent in terms of computing power. NFAs and DFAs recognize
exactly the regular languages.

%% \todo[inline]{jan: What about "Nondeterministic finite automata with
%%   epsilon-transitions (FND-epsilon or epsilon-NFA)"}


\begin{description}
\item[NFA] For each pair of input symbol and state, there may be more
  next states. This means that there may be several paths through an
  NFA for a given input string.

  The $\upvarepsilon$-transitions are an extension of the NFA. These
  are special transitions that can be taken without consuming any
  input symbols. This also has mainly practical implications, NFAs
  with and without $\upvarepsilon$-transitions are equivalent.

\item[DFA] For each pair of input symbol and state, there may be only
  one next state. This means there is only one path through the DFA
  for a given input string.
\end{description}

The advantage of a NFA is the size, the number of states and
transitions in a NFA is linear in the size of the regular expression,
whereas a DFA will in the worst case have a number of states
exponential in the size of the regular expression. The advantage of a
DFA is the effort required to simulate it. It only requires time
linear in the size of the input string and constant space (not
counting the space for the DFA), whereas the NFA requires time linear
in the product of the size of the regular expression and input string
and space linear in the size of the regular expression.

%% \todo[inline]{jan: Aren't you forgetting the size growth of either?
%%   Also you're not talking about the properties of either, and their
%%   respective disadvantages}


\input{parsing}

\subsection{Summary}

Regular expressions are a widely used and popular tool. The features
offered and the semantics vary. For example some will offer back
referencing and others will not, some will offer a leftmost match in
alternations, others will offer a longest match. Even for engines with
similar feature sets, the underlying implementation and performance
can vary widely. A regular expression engine can typically solve some
types of problems more efficiently than others, or vice versa: it may
be particularly bad at a given problem. %\todo{Jan: changed}

There are many highly specialized regular expression engines
exemplifying this. To briefly mention an example: Structured text like
SGML documents benefits from a different approach. Many times you will
need to find the text between two tags, but many tools are not geared
for this kind of search: The search can span several lines and we will
usually want a shortest match. See \cite{pedersen2010} for details.

%\todo[inline]{Might want to give an example}

%% \todo[inline]{write something about how regular expressions can be
%%   very powerful but the underlying design and its performance can vary
%% widely}
